<span style="color:red;">Building a house with<br />Advanced Normalization Tools (*ANTs*)</span>

### 2014-09-03 ants2015build.Rmd
WIP: regular update

>>>>

# Registration & statistics:<br /> Frontiers and innovation

## multivariate statistical fields arise from fused modalities

![images at measurement fields](figures/statisticalfields.png)

*Many opportunities for statistical advancements*

## ITK+ANTs+R = <span style="color:red;">*ANTsR*</span>

## Agnostic statistics
![](./figures/antsrex.jpg)

[McMillan et al.](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3961542/)

## A Quick <span style="color:grey;">*ANTsR*</span> example
```{r loadantsr,results='hide',echo=FALSE,message=FALSE}
library(ANTsR)
```

This is an executable *ANTsR* code block - *N*-dimensional statistics to go with 
our *N*-dimensional image processing software!
```{r loadimages,message=FALSE,warning=FALSE,results='hide'}
library(ANTsR)
dim<-2
filename<-getANTsRData('r16')
img<-antsImageRead( filename , dim )
filename<-getANTsRData('r64')
img2<-antsImageRead( filename , dim )
mask<-getMask(img,50,max(img),T)
mask2<-getMask(img,150,max(img),T)
nvox<-sum( mask == 1 )
nvox2<-sum( mask2 == 1 )
```
The brain has `r nvox` voxels ...

## A Quick <span style="color:grey;">*ANTsR*</span> example

Simulate a population morphometry study - a "VBM" ...
```{r morph,results='hide'}
simnum<-10
imglist<-list()
imglist2<-list()
for ( i in 1:simnum ) {
  img1sim<-antsImageClone(img)
  img1sim[ mask==1 ]<-rnorm(nvox,mean=0.5)
  img1sim[ mask2==1 ]<-rnorm(nvox2,mean=2.0)
  img2sim<-antsImageClone(img2)
  img2sim[ mask==1 ]<-rnorm(nvox,mean=0.20)
  imglist<-lappend(imglist,img1sim)
  imglist2<-lappend(imglist2,img2sim)
}
imglist<-lappend( imglist, imglist2 )
mat<-imageListToMatrix( imglist, mask )
DX<-factor( c( rep(0,simnum), rep(1,simnum) ) )
mylmresults<-bigLMStats( lm( mat ~ DX ) )
qvals<-p.adjust( mylmresults$pval.model ) 
```
The minimum q-value is `r min(qvals)` ...

## Visualize the histograms of effects
```{r vizmorph,results='hide'}
whichvox<-qvals < 1.e-2
voxdf<-data.frame( volume=c( as.numeric( mat[,whichvox] ) ), DX=DX )
ggplot(voxdf, aes(volume, fill = DX)) + geom_density(alpha = 0.2)
```

## Visualize the anatomical distribution
```{r vizmorph2hide,results='hide',echo=FALSE}
betas<-antsImageClone(mask)
betavec<-abs(mylmresults$beta.t[1,])
betas[mask==1]<-betavec
SmoothImage(dim,betas,1.5,betas)
thresh<-paste(3.0,'x',max(betavec),sep='')
ofn<-paste(figpath,'vizmorph2.jpg',sep='')
```

```{r vizmorph2,results='hide'}
plotANTsImage(img,functional=list(betas),threshold=thresh,
  outname=ofn)
```

![](`r ofn`)

## Network visualization
see `?plotBasicNetwork`
```{r networkfn,results='hide',warning=FALSE,echo=FALSE,message=FALSE}
ofn<-paste(figpath,'network',sep='')
ofnpng<-paste(ofn,'.png',sep='')
```
```{r network,results='hide',warning=FALSE,echo=FALSE,message=FALSE,eval=T}
       mnit<-getANTsRData("mni")
       mnit<-antsImageRead(mnit,3)
       mnia<-getANTsRData("mnia")
       mnia<-antsImageRead(mnia,3)
       ThresholdImage(3,mnit,mnit,1,max(mnit))
       ImageMath(3,mnit,"FillHoles",mnit)
       cnt<-getCentroids( mnia, clustparam = 50 )
       aalcnt<-cnt$centroids[1:90,]
       brain<-renderSurfaceFunction( surfimg =list( mnit ) , alphasurf=0.1 ,smoothsval = 1.5 )
       testweights<-matrix( rep( 0, 90*90 ) ,nrow=90)
       testweights[31,37]<-1  # ant cingulate to hipp
       testweights[31,36]<-2  # ant cingulate to post cingulate
       testweights[11,65]<-3  # broca to angular 
       plotBasicNetwork( centroids = aalcnt , brain , weights=testweights )
       id<-par3d("userMatrix")
       rid<-rotate3d( id , -pi/2, 1, 0, 0 )
       rid2<-rotate3d( id , pi/2, 0, 0, 1 )
       rid3<-rotate3d( id , -pi/2, 0, 0, 1 )
       par3d(userMatrix = id ) 
       dd<-make3ViewPNG(  rid, id, rid2,  ofn )
       par3d(userMatrix = id ) 
```
![](`r ofnpng`)

## The power of *ANTs* $+$ *R* $\rightarrow$<br /> <span style="color:red;">**Reproducible imaging science**</span>

![](./figures/antsgoals.png)

... used in ["Sparse canonical correlation analysis relates network-level atrophy to multivariate cognitive measures in a neurodegenerative population"](http://www.ncbi.nlm.nih.gov/pubmed/24096125) and several upcoming ...
